---
title: Data
description: We describe the sources of our data and the cleaning process.
toc: true
draft: false
---
# NYPD Shooting Data - Team 6
  
## Data Background

### NYPD Shooting Incident Data 

This dataset encompasses a detailed account of each shooting incident in NYC from 2006 up until the end of the last calendar year. Compiled and meticulously reviewed quarterly by the Office of Management Analysis and Planning, it is then made accessible on the NYPD website. Every entry in this dataset chronicles a specific shooting event, detailing the incident's circumstances, location, and timing. Additionally, it provides demographics regarding suspects and victims.

The data set primarily focuses on shooting incidents within New York, so it inherently limits the possibility of generalizing our findings to other locations. The age is categorized by ranges, such as 18-24, 25-44, and 45-64, which may reduce the precision of our analysis. Without a more detailed record of the shooting incidents, we do not know the number of perpetrators and victims. The lack of a detailed understanding of the level of severity of the incidents may cause bias in our analysis.

We are able to download the dataset from the original source. Link to the data:
[NYPD Police Shootings](https://catalog.data.gov/dataset/nypd-shooting-incident-data-historic)

### NYPD Arrest Data (Combined Data)

We combined the two datasets based on Latitude and Longitude to explore potential relationships between shooting incidents across each borough in NYC. This dataset includes the offense type and the policy description for each shooting incident. By merging these datasets, we can analyze comprehensive details surrounding each incident, including the types of offenses and the corresponding police actions. This approach not only enhances the depth of the data but also supports a more robust statistical analysis to better understand the dynamics of crime and law enforcement responses in New York City.

### NYC Payroll Data

This dataset gives the information of the average basic salary for each borough in NYC. Incorporating the dataset detailing the average basic salary for each borough in NYC into our analysis offers a valuable dimension to understanding the socio-economic context of shooting incidents. By examining the relationship between economic factors and crime rates across different boroughs, we can explore whether disparities in income levels correlate with variations in shooting incidents. This combined approach allows us to investigate not just the geographic distribution of crime, but also how economic conditions might influence criminal activities.

## Summary of the Dataset structure and some key statistics

We will concentrate our analysis on examining potential differences among NYC boroughs in shooting crime incidents, focusing specifically on the interplay between the perpetrator's and victim's information, including their race, age, sex, in conjunction with offense types and other possible crucial factors. 

- **Location Data**: There are multiple columns related to the location of the incidents, such as BORO, PRECINCT, X_COORD_CD, Y_COORD_CD, Latitude, and Longitude.

- **Incident Details**: Includes details like OFNS_DESC, LOC_CLASSFCTN_DESC (Location Classification Description), and LOCATION_DESC.

- **Perpetrator and Victim Information**: Information on the perpetrator (PERP_AGE_GROUP, PERP_SEX, PERP_RACE) and victim (VIC_AGE_GROUP, VIC_SEX, VIC_RACE) is available.

- **Statistical Flags**: Includes a flag for statistical murders (STATISTICAL_MURDER_FLAG)

## Data Loading and Cleaning

```{r message=FALSE, warning=FALSE, results='hide'}
source(
  "scripts/load_and_clean_data.R", echo = FALSE)
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)

# Read the dataset
NYPD_data <- read_csv("dataset/NYPD_Shooting_Incident_Data__Historic__20240322.csv")

NYPD_clean_data <- NYPD_data %>%
  select(-c(OCCUR_TIME, LOC_OF_OCCUR_DESC)) %>%
  filter(!is.na(PERP_RACE) & !is.na(VIC_RACE)) %>%
  mutate(OCCUR_DATE = dmy(OCCUR_DATE), 
         OCCUR_YEAR = year(OCCUR_DATE))
```
### Cleaning Process

- We removed the OCCUR_TIME and LOC_OF_OCCUR_DESC as we did not find them useful for our analysis.
- We changed  'OCCUR_DATE' to DateTime format.
- We also removed columns where we did not have the perpetrator's race or the victim's race.

## Combining Datasets

```{r}
NYPD_arrest_data <- readRDS("dataset/nypd_arrest.rds")
combined_df <- merge(NYPD_clean_data, NYPD_arrest_data, by = c("X_COORD_CD", "Y_COORD_CD","Latitude","Longitude","PERP_SEX","JURISDICTION_CODE","PERP_RACE"))
```

### Other Dataset Used in the Project Analysis

```{r}
nypd_payroll <- readRDS("dataset/nypd_payroll.rds")
```