---
title: "Blog Post 4 - EDA"
author: "Team"
date: "2024-04-08"
date-modified: "2024-04-08"
draft: FALSE
---
```{r}
# Load necessary libraries
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(ggcorrplot))
suppressPackageStartupMessages(library(GGally))
suppressPackageStartupMessages(library(readr))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(lubridate))

# Read the dataset
data <- readRDS("dataset/nypd_shooting_clean.rds")
```

```{r}
# Convert 'OCCUR_DATE' to datetime format
data <- data %>%
  mutate(OCCUR_DATE = dmy(OCCUR_DATE), 
         OCCUR_YEAR = year(OCCUR_DATE))
```

```{r}
prep_race_counts <- data |>
  count(PERP_RACE) |>
  arrange(desc(n))

ggplot(prep_race_counts, aes(x = reorder(PERP_RACE, n), y = n, fill = PERP_RACE)) + 
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Perpetrator Race Distribution", x = "Perpetrator Race", y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

We created a bar chart to analyze the perpetrator’s race, a key variable of interest for later investigation in our project. The chart revealed approximately 600 missing (null) values for this variable, indicating a potential challenge that requires our attention.

```{r}
# Calculate proportions Blog Post 4
proportions <- data %>%
  group_by(PERP_RACE) %>%
  summarise(Proportion = mean(STATISTICAL_MURDER_FLAG == "TRUE", na.rm = TRUE))

# Create the plot
ggplot(proportions, aes(x = PERP_RACE, y = Proportion, fill = PERP_RACE)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Proportion of Incidents Classified as Statistical Murders by Perpetrator Race",
       x = "Perpetrator Race",
       y = "Proportion") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
mod1 <- glm(
  STATISTICAL_MURDER_FLAG ~ PERP_RACE + PERP_AGE_GROUP + PERP_SEX,
  data = data, 
  family = "binomial"
)
mod1
```

We made a logistic model using murder statistics as y variable and PERP_RACE, PERP_AGE_GROUP, and PERP_SEX as predictors. And based on the bar plot above, Asian/pacific islander has the largest proportion of Incidents Classified as Statistical Murders by Perpetrator Race.